Supplemental Information: Aligning marine species range data to better serve science and conservation

### 2016-09-12 ms_0_abst.Rmd
argh this fucking abstract

The Abstract should:
  Describe the main objective(s) of the study
  Explain how the study was done, including any model organisms used, without methodological detail
  Summarize the most important results and their significance
  Not exceed 300 words
Abstracts should not include:
  Citations
  Abbreviations, if possible
-->

Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and inconsistencies.  AquaMaps and the International Union for Conservation of Nature (IUCN) are two distinct efforts to map marine species distributions at a global scale.  Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.  Here we examine differences in predicted species ranges between the two datasets and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.  We illustrate the scientific and management implications of these differences by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.  While we suggest methods to address specific errors in each spatial dataset, it remains essential to understand the implications of dataset differences for conservation planning and decision-making.

>>>>

The Abstract should:
  Describe the main objective(s) of the study
  Explain how the study was done, including any model organisms used, without methodological detail
  Summarize the most important results and their significance
  Not exceed 300 words
Abstracts should not include:
  Citations
  Abbreviations, if possible
-->
from intro:
Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation.  This body of science has resulted in various compiled databases of species distribution and range maps which provide foundational information for understanding species diversity and extinction risk, predicting species responses to human impacts and climate change, and managing and protecting species effectively.  No species range dataset can claim to represent the "truth" of any species' spatial distribution, but rather each offers its own distinct understanding of that latent distribution.  These varying predictions of species presence and absence, driven by intent and methodology, can result in very different predictions of species presence; however, conservation managers and policy makers must base their actions on conclusions drawn from these imperfect datasets.  It is critical to understand the differences in spatial range datasets and the implications of these differences for conservation research and decision-making.

Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions.  Two major efforts to provide marine species distributions at a global scale are the International Union for Conservation of Nature (IUCN), which provides expert-generated range maps that outline the complete extent of a species' distribution; and AquaMaps, which provides model-generated species distribution maps that predict areas occupied by the species.  Neither dataset claims to represent the "truth" of species' spatial distributions, but rather each communicates a distinct understanding of the underlying distribution, with a distinct purpose in mind.

Distinctions in intent and methodology can result in very different predictions of species distributions; these differences bear important implications for scientists and decision makers who rely on these datasets to inform conservation policy and management actions.

Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.  Ideally, the decision to use one dataset one over the other should carefully consider the purpose for which it was designed, but this small overlap between the two suggests that data availability is likely a more important factor.

Comparing these overlapping predicted species ranges, we highlight several important differences between the two datasets, tradeoffs, blah blah

We illustrate the scientific and management implications of these differences/tradeoffs by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.

While we suggest methods to address specific errors in each spatial dataset, it remains essential to understand the implications of dataset differences for conservation planning and decision-making.

FRAME as tradeoffs rather than differences; this could be a way to bring the two datasets together for the last paragraph of the results and discussion - minimizing tradeoffs rather than fixing differences?

### 2016-09-12 ms_1_intro.Rmd
argh this fucking abstract

While the two datasets ostensibly describe the same information, i.e. where can a particular species be found, they communicate fundamentally distinct concepts.  The IUCN range map for a given species describes the "extent of occurrence" for that species, a contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].  The AquaMaps map for a species, on the other hand, seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found, for example by excluding unsuitable habitats from the extent of occurrence [@iucn_iucn_2015].  The distinctions between extent of occurrence and area of occupancy, including methodology, limitations, and appropriate usage, are well documented [@rondinini_tradeoffs_2006; @gaston_sizes_2009]. ___include reference to a supplemental table for more details?___

@@

Data availability is a crucial factor in deciding which dataset to use.  We compare how each data source represents the global spatial and taxonomic distribution of the `r format(n_spp_all, big.mark = ',')` marine species mapped by one or both datasets.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).  Yet the small overlap in mapped species between the two datasets means each provides unique value in taxonomic and geographic coverage.

Although relatively small in number, these overlap species present a unique opportunity to evaluate the two datasets overall.  For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both spatial distribution and overall area.  Based upon the definitions of extent of occurrence and area of occupancy, we expect that for a given species, the AquaMaps predicted distribution will fall within, and describe a smaller range than, the IUCN predicted distribution.  Where these expectations seem to fail, we explore methodological issues that can introduce errors.  In particular, we show how explicitly using depth as a constraint on coral ranges in IUCN range data dramatically mitigates a source of commission error, and highlight a side effect of AquaMaps methodology that results in discontinuous species range boundaries.

>>>>

While the two datasets ostensibly describe the same information, i.e. where can a particular species be found, they communicate fundamentally distinct concepts.  The IUCN range map for a given species describes the "extent of occurrence" for that species, a contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that [the species] is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].  The AquaMaps map for a species, on the other hand, seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found, for example by excluding unsuitable habitats from the extent of occurrence [@iucn_iucn_2015].  The distinctions between extent of occurrence and area of occupancy, including methodology, limitations, and appropriate usage, are well documented [@rondinini_tradeoffs_2006; @gaston_sizes_2009]. ___include reference to a supplemental table for more details?___

@@

Data availability is a crucial factor in deciding which dataset to use.  We compare how each data source represents the global spatial and taxonomic distribution of the `r format(n_spp_all, big.mark = ',')` marine species mapped by one or both datasets.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps.  IUCN range map data exist for `r format(n_spp_iucn, big.mark = ',')` unique marine species, with only `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets.  This small overlap means each dataset provides unique value in taxonomic and geographic coverage, but also suggests that data availability likely plays a larger role in selecting a dataset than does the intended purpose of the data.

Although relatively small in number, these overlap species present a unique opportunity to evaluate the two datasets overall.  For the species included in both datasets, we examine how well the maps align in both spatial distribution and overall area.  Based upon the definitions of extent of occurrence and area of occupancy, we expect that for a given species, the AquaMaps predicted distribution will fall within, and describe a smaller range than, the IUCN predicted distribution.  Where these expectations seem to fail, we explore methodological issues that can introduce errors.  In particular, we show how explicitly using depth as a constraint on coral ranges in IUCN range data dramatically mitigates a source of commission error, and highlight a side effect of AquaMaps methodology that results in discontinuous species range boundaries.
